---
title: "ml_project"
output: html_document
---

# Machine learning Project

## Data

data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

## Goals

The goal of this project is to predict the manner in which they did the exercise. This is the **classe** variable in the training set.

### data

1. load data and libraries

```{r}
train <- read.csv("./train.csv")
validation <- read.csv("./test.csv")
library(caret)
library(ggplot2)
library(parallel)
library(doParallel)
```

2. partition data

```{r}
set.seed(369)
inTrain <- createDataPartition(train$classe, p = .8, list = FALSE)
training <- train[inTrain,]
testing <- train[-inTrain,]
```

### preprocessing

1. remove useless variables: vars that do not influence the reusult (the first 7)

```{r}
training <- training[,-c(1:7)]
testing <- testing[,-c(1:7)]
```

2. remove zero covariates

```{r}
nsv <- nearZeroVar(training, saveMetrics = TRUE)
x <- which(nsv$nzv == TRUE)
training <- training[,-c(x)]
testing <- testing[,-c(x)]
```

3. remove mostly Nas 

```{r}
mostNAs <- sapply(training, function(x) mean(is.na(x))) > 0.95
training <- training[,mostNAs == FALSE]
testing <- testing[,mostNAs == FALSE]
```

## test models

### random forest

1. configure parallel process

```{r}
cluster <- makeCluster(detectCores() - 1) # convention to leave 1 core for OS
registerDoParallel(cluster)
```

2. create fit control

```{r}
fitControl <- trainControl(method="cv", number = 4, allowParallel = TRUE)
```

3. fit model

```{r}
fit1 <- train(classe ~., method = "rf",data = training,trControl = fitControl, preProcess = c("center", "scale"))
```

4. stop parallel processing

```{r}
stopCluster(cluster)
registerDoSEQ()
```

5. check accuracy

```{r}
pred1 <- predict(fit1, testing)
confusionMatrix(pred1, testing$classe)$overall[1]
```

### boosted trees

1. fit model

```{r}
fit2 <- train(classe ~., method = "gbm",data = training,trControl = fitControl, preProcess = c("center", "scale"), verbose = FALSE)
```

2. check accuracy

```{r}
pred2 <- predict(fit2, testing)
confusionMatrix(pred2, testing$classe)$overall[1]
```

the most accurate model is the first one: random forest


## predict **classe** on test set

```{r}
predict(fit1, validation)
```

these are the associated classes to test data
